---
title: "giving-doing-paper"
subtitle: "raw code & analysis"
author: "Raph Mondesir"
date: "Sys.Date()"
output: html_document
---

# The original file used to produce the chapter 1 in my dissertation is: "W1.Models_pt2_March2017" in the folder "Chap1_Giving_Doing" in "CLEAN_DATA" in PNPM folder on dropbox.

```{r setup, include=FALSE}
## install packages if necessary
list.of.packages <- c("censReg", "texreg", "coefplot", "ggplot2", "readr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

## load packages
library(censReg)
library(texreg)
# install.packages("coefplot2",repos="http://r-forge.r-project.org")
library(coefplot)
library(ggplot2)
library(readr)

# a shorthand function for tabulations.
ftab <- function(x){
  a <- table(x, useNA = "always")
  return (a)
}

#  get rid of scientific notation
options(scipen = 999)

```

### Load Data

```{r load-data, warning=FALSE} 
#READING IN the wave 1 DATA
library(foreign)
setwd("/cloud/project/giving-doing-paper/data/raw-data")
W1.data <- read.dta("GW1-NOcentrd-0217.dta")
#  this is the data set used to produce chapter 1 (Beard replic)... even though I cleaned the data for wave I further and added new variables. this file above should not be changed in order to ensure reproducibility. OK
#
dim(W1.data)
# 22294   71
#
```

### Recode a few variables

```{r recode-1, warning=FALSE, echo=FALSE}
library(skimr)
library(tidyr)

skim(W1.data)

#RECODING RELIGION DIFFERENTLY
#-------------------------------
W1.data$RELIG2 <- "Muslim"
W1.data$RELIG2[W1.data$ReligionsRaw==2 | W1.data$ReligionsRaw==3] <- "Christian"
W1.data$RELIG2[W1.data$ReligionsRaw==4 | W1.data$ReligionsRaw==5]<- "Buddhist/Hindu"
# W1.data$RELIG2[W1.data$ReligionsRaw==5] <- "Hindu"
W1.data$RELIG2[W1.data$ReligionsRaw > 5]<- "Other"
#
ftab(W1.data$RELIG2)
#
W1.data$RELIG2 <- as.factor(W1.data$RELIG2)
W1.data$RELIG2 = factor(W1.data$RELIG2, levels=c("Muslim", "Christian", "Other", "Buddhist/Hindu"))


# centering with 'scale()' function inside a wrapper function
center_scale <- function(x) {
  scale(x, scale = FALSE)
}
#
W1.data$Assets_ctrd  <- center_scale(W1.data$TotAssets)
W1.data$KnowLeaders_ctrd  <- center_scale(W1.data$KnowLeaders)
W1.data$PctIslam_ctrd  <- center_scale(W1.data$PctIslam)


## create a new binary var ==> MajorMuslim (1/0)
W1.data$MajorMuslim <- 0
W1.data$MajorMuslim[W1.data$PctIslam > 50] <- 1
ftab(W1.data$MajorMuslim)


#  recoding "Major ethnic" as Ethnic minority ===> for easier interpretation in the models
#
# new var
W1.data$EthnicMinor <- NA
W1.data$EthnicMinor[W1.data$MajorEth==0] <- 1
W1.data$EthnicMinor[W1.data$MajorEth!=0] <- 0
ftab(W1.data$EthnicMinor)
ftab(W1.data$MajorEth)
#  ok great!
```

### Save Raw Data

Lets' save the data sets into an object in `R`.

```{r save-data, warning=FALSE} 
## use saveRDS() to save each object as a .rds file 
saveRDS(W1.data, file = "w1-data.rds") 

```

### Run the regression models

```{r model-1, warning=FALSE}

# ================== TOBIT simplified ==================
library(censReg)
cens.Doing <- censReg(TotTime ~
                        # @@@@@ Resources @@@@@
                      + PrimEduc
                      + SeniorHigh
                      + Univ
                      + factor(percapita)
                      + TotAssets
                      + Hdemployed
                      # @@@@@ NEED @@@@@
                      + HDNoEduc
                      + DirectCash
                      #
                      # @@@@@@@@@@ RELIGION @@@@@@@@@@
                      + nonMuslim
                      + nonMuslim:TotAssets
                      # @@@@@ Community Characteristics @@@@@
                      + MajorMuslim
                      + log(VilPop)
                      + homog
                      # @@@@@ Control variables @@@@@@@
                      + femhead
                      + HDMigrated
                      + MajorEth,
                  left=0, right=Inf, data = W1.data)
#
library(texreg)
# screenreg(cens.Doing, digits=3, single.row = TRUE, custom.model.names = c("Doing"))
htmlreg(cens.Doing, digits=3, single.row = TRUE, custom.model.names = c("Doing"))

```

```{r model-2, warning=FALSE}
#
cens.Giving <- censReg( # @@@@@ Household SES @@@@@
  TotMoney ~ # @@@@@ Resources @@@@@
    + PrimEduc
  + SeniorHigh
  + Univ
  + factor(percapita)
  + TotAssets
  + Hdemployed
  # @@@@@ NEED @@@@@
  + HDNoEduc
  + DirectCash
  #
  # @@@@@@@@@@ RELIGION @@@@@@@@@@
  + nonMuslim
  + nonMuslim:TotAssets
  # @@@@@ Community Characteristics @@@@@
  + MajorMuslim
  + log(VilPop)
  + homog
  # @@@@@ Control variables @@@@@@@
  + femhead
  + HDMigrated
  + MajorEth,
  left=0, right=Inf, data = W1.data)
#
# screenreg(cens.Giving, digits=3, single.row = TRUE, custom.model.names = c("Giving"))
htmlreg(cens.Giving, digits=3, single.row = TRUE, custom.model.names = c("Giving"))

```

### Save the regression models as objects that can easily be loaded next time using `readRDS()`

```{r save-models, warning=FALSE} 
saveRDS(cens.Giving, "giving-model.rds")
saveRDS(cens.Doing, "doing-model.rds")
```

# Extracting coefs and Std Errors

```{r coefs-stds, warning=FALSE} 

# 1) for DOING
#=========================
coefs.T <- coef(summary(cens.Doing))[, 1]
names(coefs.T)
se.T <- coef(summary(cens.Doing))[, 2]
names(se.T)
#  drop logsigma
se.T <- se.T[-c(1,19)]
names(se.T)
coefs.T <- coefs.T[-c(1,19)]
names(coefs.T)
#  ok

#  NOW FOR MOney model
coefs.M <- coef(summary(cens.Giving))[, 1]
names(coefs.M)
se.M <- coef(summary(cens.Giving))[, 2]
names(se.M)
#  drop logsigma
se.M <- se.M[-c(1,19)]
names(se.M)
coefs.M <- coefs.M[-c(1,19)]
names(coefs.M)
```

# Plots for the full model

```{r plots-full-1, warning=FALSE} 

longnames=c("Primary school", "Senior high", "University degree",
                      "Low percapita expenditures", "High percapita expenditures",
                      "Household Assets", "Employed head",
                      "No Education",   "Direct cash",
                      "Non-Muslim","Household Assets * Non-Muslim",
                      "Majority Muslim", "Ethnic homogeneity", "Population size (logged)",
                      "Female head","Recently migrated","Belongs to ethnic majority")


# This is a treasure!---v
par(mfrow=c(1,1), family="serif") #  using Times New Roman
#
# 
require(coefplot)
coefplot::coefplot(cens.Doing,
                     # coefs.T, se.T,
                     intercept=FALSE,
                     xlim=c(-1, 0.8),
                     v.axis=TRUE,top.axis=T,
                     #  h.axis=T,
                     cex.var=0.7, pch.pts=23, lwd.1=1, lwd.2=0.6,
                     # varnames=longnames,
                     # varnames=c("Primary school", "Senior high", "University degree",
                     #            "Low percapita expenditures", "High percapita expenditures",
                     #            "Household Assets", "Employed head",
                     #            "No Education",   "Direct cash",
                     #            "Non-Muslim","Household Assets * Non-Muslim",
                     #            "Majority Muslim", "Ethnic homogeneity", "Population size (logged)",
                     #            "Female head","Recently migrated","Belongs to ethnic majority"),
                     cex.axis=0.2,
                     mar=c(1.5, 11, 5.7,1), cex.main=0.7,
                     main="

                     Tobit regression (DOING)\n



                     ")
# the spacing and empty line are important to keep the graph elegant!
```


```{r plots-full-2, warning=FALSE}
#  calling the second plot
coefplot::coefplot(cens.Giving,
                     # coefs.M, se.M,
                     intercept=FALSE,
                     xlim=c(-2.5, 2.8),
                     v.axis=TRUE,top.axis=T,
                     #  h.axis=T,
                     cex.var=0.7, pch.pts=22, lwd.1=1, lwd.2=0.6,
                     # varnames=longnames,
                     # varnames=c("Primary school", "Senior high", "University degree",
                     #            "Low percapita expenditures", "High percapita expenditures",
                     #            "Household Assets", "Employed head",
                     #            "No Education",   "Direct cash",
                     #            "Non-Muslim","Household Assets * Non-Muslim",
                     #            "Majority Muslim", "Ethnic homogeneity", "Population size (logged)",
                     #            "Female head","Recently migrated","Belongs to ethnic majority"),
                     cex.axis=0.2,
                     mar=c(1.5, 11, 5.7,1), cex.main=0.7,
                     main="

                     Tobit regression (GIVING)\n



                     ")








```



# BETTER PLOTS FOR PUBLISHING:

```{r plots-sub, warning=FALSE} 

## TAKE A SUBSET OF THE COEFS/VARS

names(coefs.T)
#
coefs.T_sub <- coefs.T[c("HDNoEduc","factor(percapita)1", "DirectCash", "TotAssets", "Hdemployed",
                         "Univ","factor(percapita)3")]
names(coefs.T_sub)
#
se.T_sub <- se.T[c("HDNoEduc","factor(percapita)1", "DirectCash", "TotAssets", "Hdemployed",
                   "Univ","factor(percapita)3")]
names(se.T_sub)

#

names(coefs.M)
#
coefs.M_sub <- coefs.M[c("HDNoEduc","factor(percapita)1", "DirectCash", "TotAssets", "Hdemployed",
                           "Univ","factor(percapita)3")]
names(coefs.M_sub)
#
se.M_sub <- se.M[c("HDNoEduc","factor(percapita)1", "DirectCash", "TotAssets", "Hdemployed",
                     "Univ","factor(percapita)3")]
names(se.M_sub)
#

par(mfrow=c(1,1), family="serif", col.main = "black") #  using Times New Roman

#
#
# DOING
# png(filename = "/giving-doing-paper/figures/explanatory-figures/coefplots-bw.png", width=1600, height=700) # this was for the plot in black and white. where the models differed by the shape for coeffs.
#

# colored coeff plots.
png(filename = "/giving-doing-paper/figures/explanatory-figures/coefplots-col.png",
    width=1600, height=700)

par(mfrow=c(1,2),  family="serif", col.main = "black")
#
gr.T <- coefplot(coefs.T_sub, se.T_sub, xlim=c(-2, 2.2),
                             # xlim=c(-0.7, 0.8),
                     v.axis=TRUE,top.axis=T,
                     # h.axis=T,
                     cex.var=2.4, lwd.1=1.8, lwd.2=1.1,
                     pch.pts=20, cex.pts=5.1, col="#98777b",
                     varnames=c("No education",
                                "Low per capita expenditures",
                                "Direct cash assistance",
                                #
                                "Household assets",
                                "Employed, head",
                                "University degree",
                                "High percapita expenditures"),
                     cex.axis=18,
                     mar=c(2.5, 27, 7.8,1),
                     cex.main=2.5,
                     # legend=T, legend.arg=legend("bottomright", "Giving: Red\n Doing: Blue", col=c("tomato", "blue")),
                     main="Doing (Y = Time)\n")
# (controlling for social capital, ethnic homogeneity, gender, \nmigration, ethnicity, vilage population size, and household size)

#
#
# GIVING
gr.M <- coefplot(coefs.M_sub, se.M_sub, xlim=c(-2, 2.2),
                     v.axis=TRUE,top.axis=T,
                     # h.axis=T,
                     cex.var=2.4, lwd.1=1.8, lwd.2=1.1,
                      # pch.pts=11,
                     pch.pts=20,
                     cex.pts=3.5, col="blue",
                     varnames=c("No education",
                                "Low per capita expenditures",
                                "Direct cash assistance",
                                #
                                "Household assets",
                                "Employed, head",
                                "University degree",
                                "High percapita expenditures"),
                     cex.axis=16,
                     mar=c(2.5, 28, 7.8,1),
                     cex.main=2.5,
                     # legend=T, legend.arg=legend("bottomright", "Giving: Red\n Doing: Blue", col=c("tomato", "blue")),
                     main="Giving (Y = Money)\n")
# (controlling for social capital, ethnic homogeneity, gender, \nmigration, ethnicity, vilage population size, and household size)
#
dev.off()

# $$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$$
#
```

### this file won't need because of the proble I describe below:
### the `coefplot` graphs did not really work here, but they worked in the lcoal workspace on my laptop/dropbox. That's ok. For some reason, the `coefplot2` package has been maintained/updated to work with the later versions of R. too bad, it was pretty neat. Ben Bolker should tend to it! 

### End 

